\documentclass[11pt]{letter}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage{geometry}
\geometry{margin=1in}

\signature{Haobo Ma and Wenlin Zhang}
\address{Wenlin Zhang (Corresponding Author)\\National University of Singapore\\Singapore\\e1327962@u.nus.edu}
\date{\today}

\begin{document}
\begin{letter}{Editorial Board\\Nonlinearity}

\opening{Dear Editors,}

We submit our manuscript entitled

\medskip
\centerline{\textbf{A sharp three-window threshold and finite-memory}}
\centerline{\textbf{conjugacy in Fibonacci stabilization}}
\medskip

\noindent for consideration in \textit{Nonlinearity}.

The paper proves a sharp overlap threshold for the canonical Fibonacci stabilization map $\mathrm{Fold}_m:\{0,1\}^m\to X_m$, which sends a binary word to the unique admissible Fibonacci word with the same valuation modulo $F_{m+2}$. The main contributions are:

\begin{itemize}
\item A sharp three-window threshold: for every $m\ge 3$, the sliding stabilized-window map is injective and admits an explicit finite-memory inverse, identifying the image with a subshift of finite type topologically conjugate to the full two-shift. The constructive decoder at $m=3$ is given by explicit modular arithmetic.

\item A complete classification of the exceptional case $m=2$: injectivity fails on exactly one branch orbit, with exact conditional entropy and Kullback--Leibler defect formulas. Off this branch set the map is a measure-theoretic isomorphism.

\item Above threshold, the conjugacy transports invariant measures, continuous potentials, pressure, and equilibrium states exactly. Parameterized phase diagrams, finite-range Hamiltonians, and block-level information geometry are all preserved.

\item In the Sturmian benchmark, quantitative discrepancy bounds control stabilized block laws and the recovery of finite-range observables, with explicit error terms involving the star discrepancy and boundary-layer mass.
\end{itemize}

We believe the manuscript fits the scope of \textit{Nonlinearity} because it establishes a sharp structural phenomenon in symbolic dynamics---the arithmetic threshold for finite-memory invertibility of Fibonacci stabilization---with exact consequences for thermodynamic formalism. The interplay between Fibonacci residue arithmetic and sofic-factor theory produces results (the threshold decoder, the branch-locus classification, the affine transfer principle) that are specific to this number-theoretic setting and do not follow from general sofic-shift frameworks.

The manuscript is original, has not been published previously, and is not under consideration elsewhere. All authors have approved the submission.

\closing{Sincerely,}

\end{letter}
\end{document}
