\section{Introduction}\label{sec:intro}

Finite-window recodings arise naturally in symbolic dynamics, numeration, and finite-resolution observation. A local map may be many-to-one on isolated words and yet become invertible once one passes to overlapping windows along a trajectory. The problem studied here is whether this recovery-by-overlap occurs for the canonical Fibonacci stabilization map, and if so at what window size.

For each $m\ge 1$ we consider
\[
\Fold_m:\{0,1\}^m\to X_m,
\]
the map that sends a binary word to the unique Fibonacci-admissible word with the same valuation modulo $F_{m+2}$. On individual words $\Fold_m$ is genuinely many-to-one: different binary blocks can lie in the same residue class and therefore stabilize to the same admissible block. The central question is whether this local folding persists on path space for the associated sliding-block map, or whether overlap restores the lost information.

The answer is sharp. The main contributions of the paper are:
\begin{enumerate}[label=(\roman*),leftmargin=*,itemsep=2pt]
  \item The overlap graph yields a finite-state lift with bounded degree, and the threshold window is exactly $m=3$: Lemma~\ref{lem:m3-decoder} gives an explicit three-window decoder and Theorem~\ref{thm:block-invertibility} proves injectivity for every $m\ge 3$.
  \item The exceptional case $m=2$ is completely rigid rather than generically non-injective. Theorem~\ref{thm:m2-branch-locus} identifies a unique branch orbit, Corollary~\ref{cor:m2-defect-formulas} gives exact entropy and Kullback--Leibler defect formulas, and Corollary~\ref{cor:m2-mod-zero} shows mod-zero invertibility off that branch set.
  \item Above threshold, the stabilized image is a finite-type recoding of the full two-shift. Theorem~\ref{thm:two-sided-conjugacy} supplies a finite-memory inverse and identifies the image shift $Y_m$ as a subshift of finite type.
  \item The conjugacy has exact dynamical consequences. Proposition~\ref{prop:conjugacy-consequences} and Theorem~\ref{thm:continuous-potentials} transport invariant measures, entropy, continuous potentials, pressure, and equilibrium states exactly, while Section~\ref{sec:fp:tv} gives quantitative discrepancy-based corollaries in the Sturmian benchmark.
\end{enumerate}

This places the paper closer to symbolic dynamics and arithmetic coding than to a measurement-theory manuscript. The cut-and-project and detector language serve only to fix the benchmark source law and to explain why the local problem must be posed on the full cube $\{0,1\}^m$. Once the source law and admissible grammar are fixed, the proofs are symbolic and arithmetic.

The relevant literature splits into four strands. Classical work on Sturmian words and Fibonacci numeration explains admissibility, normalization, and automaton structure at the level of words \cite{MorseHedlund1940,Lothaire2002,Queffelec2010,Frougny1992,Berstel2001}; the broader interaction between numeration systems and symbolic dynamics is surveyed in \cite{BertheRigo2010}, and $S$-adic generalizations appear in \cite{BertheSteinerThuswaldner2019}. General sofic-factor theory provides the language of labeled graphs, resolving maps, and finite-to-one factors \cite{LindMarcus1995,Kitchens1998,Nasu1995}; within that framework, resolving factor maps between irreducible shifts of finite type with equal entropy are known to be conjugacies under general conditions \cite{Ashley1988}, and lower-entropy sofic factors have been characterized abstractly \cite{Boyle1983}. The connection between $\beta$-expansions and shift spaces was introduced by Parry~\cite{Parry1960}. Standard thermodynamic formalism explains what must be preserved under topological conjugacy \cite{Walters1982}. What these frameworks do not provide on their own is the arithmetic threshold phenomenon proved here: the explicit fact that $m=2$ fails in exactly one branch locus, whereas every $m\ge 3$ admits a finite-memory inverse. That conclusion depends on the residue arithmetic of Fibonacci stabilization and, in particular, on the decoder of Lemma~\ref{lem:m3-decoder}; it is not a formal consequence of the general sofic setup.

To summarize the new content explicitly: the sharp $m=3$ threshold (Theorem~\ref{thm:block-invertibility}), together with the constructive decoder that proves it (Lemma~\ref{lem:m3-decoder}); the complete branch-locus classification at $m=2$ (Theorem~\ref{thm:m2-branch-locus}) with exact entropy and Kullback--Leibler defect formulas (Corollary~\ref{cor:m2-defect-formulas}); the finite-memory conjugacy to the full two-shift (Theorem~\ref{thm:two-sided-conjugacy}); the discrete Fourier formula for fiber cardinalities (Theorem~\ref{thm:fiber-dft}); the second-moment linear recurrence (Theorem~\ref{prop:second-moment-recurrence}); and the affine transfer principle connecting canonical and shifted-weight residue models (Theorem~\ref{thm:affine-transfer}). The thermodynamic transport theorems of Section~\ref{sec:measurement} follow from the conjugacy by standard machinery, but their explicit local form in this arithmetic setting has not appeared before.

A second conceptual point is that fixed-window aliasing and path-space information loss are different objects. The entropy ledger of Section~\ref{sec:entropy-ledger} measures the unresolved multiplicity of isolated windows. Above threshold this does not become a new source of dynamical irreversibility, because Section~\ref{sec:measurement} proves that overlap reconstructs the raw trajectory exactly. The local defect at $m=2$ is therefore a sharp exceptional branch, not evidence of a generic extensive loss mechanism.

The Fibonacci benchmark is a natural place to study this question. It is the smallest aperiodic binary system in which the admissible language, residue fibers, threshold decoder, and discrepancy bounds can all be written explicitly. This makes it possible to prove the path-space consequences at theorem level rather than to suggest them heuristically.

The original motivation nevertheless comes from finite-resolution observation of quasiperiodic order \cite{Shechtman1984,Jagannathan2021,BaakeGrimm2013}. In the Sturmian benchmark, any strictly positive local detector noise makes every binary word of length $m$ appear with positive probability, so the ambient cube is physically forced at the readout layer. That application explains why the local folding problem is worth studying, but it does not determine the mathematical identity of the paper.

The paper is organized as follows. Section~\ref{sec:readout} records the symbolic setup, the Sturmian benchmark, and the full-support mechanism. Section~\ref{sec:stabilization} defines Fibonacci stabilization on words, gives the residue description of the fibers, and includes a concrete worked example. Section~\ref{sec:stab:conjugacy} develops the overlap graph, the sharp three-window threshold, and the two-window branch locus. Section~\ref{sec:entropy-ledger} isolates the fixed-window aliasing ledger, and Section~\ref{sec:measurement} turns the threshold theorem into finite-memory conjugacy, path-space equivalence, and transport of thermodynamic formalism. Sections~\ref{sec:fiber-spectrum} and~\ref{sec:fp:tv} record arithmetic fallout and quantitative discrepancy bounds in the Sturmian benchmark. Appendix~\ref{sec:superspace} contains the source-layer model-set input.
