\section{Symbolic setup and support inflation}\label{sec:readout}

The theorem-level results concern a local Fibonacci folding on the full cube and its overlap dynamics. This section records only the symbolic input imported from the source layer: the Sturmian benchmark fixes the admissible grammar, and strictly positive local noise explains why the local problem must be posed on all of $\{0,1\}^m$ rather than only on admissible words.

\subsection{Coding map and induced law}\label{sec:readout:maps}

At the symbolic layer we only need a measurable coding map from the source dynamics into a finite alphabet. The detailed detector parameters are suppressed here; they matter only through the induced symbolic law.

\begin{definition}[Finite-alphabet readout]\label{def:readout}
Fix a measurable map $R:H\to\mathcal{A}$ into a finite alphabet $\mathcal{A}$. Given a scan sequence $(h_t)_{t\in\NN}\subset H$ (typically $h_t=h_0+t\alpha$), the \emph{readout process} is the symbolic sequence
\[
(A_t)_{t\in\NN},\qquad A_t:=R(h_t)\in\mathcal{A}.
\]
\end{definition}

In the simplest case $\mathcal{A}=\{0,1\}$ and $R(h)=\mathbf{1}_W(h)$, this is a binary indicator readout. Finite-resolution variants replace the sharp indicator by a smoothed kernel that converges to $\mathbf{1}_W$ as the resolution parameter tends to zero.

\begin{definition}[Coding map and induced input law]\label{def:coding-law}
Let $(H,\mathcal{B},\mu,T)$ be the scan dynamics at the source layer, where $\mu$ is $T$-invariant. The \emph{coding map} induced by $R$ is
\[
\iota_R:H\to\mathcal{A}^{\NN},\qquad
\iota_R(h):=\bigl(R(T^t h)\bigr)_{t\ge 0}.
\]
Let $\sigma:\mathcal{A}^{\NN}\to\mathcal{A}^{\NN}$ denote the left shift. The \emph{induced input law} on symbolic sequences is the pushforward
\[
\nu_R:=(\iota_R)_*\mu.
\]
\end{definition}

\begin{proposition}[Stationarity and ergodicity of the induced law]\label{prop:induced-law}
The induced law $\nu_R$ is $\sigma$-invariant. If $\mu$ is ergodic for $T$, then $\nu_R$ is ergodic for $\sigma$.
\end{proposition}

\begin{proof}
For a cylinder $[a_0\cdots a_{m-1}]$ in $\mathcal{A}^{\NN}$,
\[
\nu_R(\sigma^{-1}[a_0\cdots a_{m-1}])
=\mu\bigl(\{h:\ R(T^{t+1}h)=a_t,\ t=0,\dots,m-1\}\bigr).
\]
By $T$-invariance of $\mu$, this equals
\[
\mu\bigl(\{h:\ R(T^{t}h)=a_t,\ t=0,\dots,m-1\}\bigr)=\nu_R([a_0\cdots a_{m-1}]),
\]
so $\nu_R$ is shift-invariant. Ergodicity is preserved under factors \cite{Walters1982}.
\end{proof}

\subsection{The Fibonacci/Sturmian benchmark}\label{sec:readout:sturmian}

When the scan dynamics is an irrational circle rotation and the readout is a two-interval coding, the resulting symbolic process is Sturmian \cite{MorseHedlund1940,Lothaire2002,Queffelec2010}. This benchmark is singled out by its minimal complexity among aperiodic binary codings and by the fact that, in the golden regime, it lands naturally in the grammar used later for stabilization.

\begin{proposition}[Rotation-to-grammar bridge in the golden regime]\label{prop:sturmian-no11}
Let $T:\TT^1\to\TT^1$ be the rotation $T(\theta)=\theta+\alpha\pmod{1}$ and take the readout
\[
R(\theta):=\mathbf{1}_{[0,\,1-\alpha)}(\theta).
\]
If $\alpha\ge \tfrac12$, then the induced symbolic process $A_t:=R(T^t\theta_0)$ is $11$-free, hence belongs to the binary subshift that forbids the word $11$. In particular this holds for the golden slope $\alpha=\varphi^{-1}$.
\end{proposition}

\begin{proof}
If $A_t=1$, then $\theta_t:=T^t\theta_0\in[0,1-\alpha)$, so $\theta_{t+1}=\theta_t+\alpha\in[\alpha,1)$. When $\alpha\ge\tfrac12$ the intervals $[0,1-\alpha)$ and $[\alpha,1)$ are disjoint, so $A_{t+1}=0$.
\end{proof}

\begin{remark}[Role of the Sturmian benchmark]
Proposition~\ref{prop:sturmian-no11} shows that a noiseless Fibonacci source lands inside the admissible language~$X_m$. Combined with Theorem~\ref{thm:fold-suite}(ii), this means that local stabilization is trivial on the benchmark source itself: $\Fold_m$ acts as the identity on every observed word. The benchmark is retained because it canonically supplies the grammar used by stabilization and because its linear word complexity drives the total-variation bound of Section~\ref{sec:fp:tv}.
\end{remark}

\subsection{Support inflation to the full cube}\label{sec:readout:noise}

The noiseless Fibonacci benchmark itself does not populate the ambient cube. The next theorem explains why the local stabilization problem must nevertheless be posed on all of $\{0,1\}^m$ once the readout layer contains any strictly positive local uncertainty.

\begin{theorem}[Non-degenerate local noise fills the ambient cube]\label{thm:noise-full-support}
Let $B=(B_t)_{t\in\NN}$ be the ideal Fibonacci/Sturmian benchmark process from Proposition~\ref{prop:sturmian-no11}, taken under its stationary rotation law. Let $A=(A_t)_{t\in\NN}$ be obtained from $B$ through a strictly positive local binary channel $Q=(Q_{ba})_{a,b\in\{0,1\}}$ in the sense that, conditional on any raw block $b_0\cdots b_{m-1}$,
\[
\PP(A_0\cdots A_{m-1}=a_0\cdots a_{m-1}\mid B_0\cdots B_{m-1}=b_0\cdots b_{m-1})
=
\prod_{t=0}^{m-1} Q_{b_t a_t},
\]
and assume $Q_{ba}>0$ for all $a,b\in\{0,1\}$. Then for every $m\ge 1$ and every word $\omega\in\{0,1\}^m$,
\[
\PP(A_0\cdots A_{m-1}=\omega)>0.
\]
In particular, the observable support of the noisy length-$m$ measurement law is the full ambient cube $\{0,1\}^m$.
\end{theorem}

\begin{proof}
Fix $m\ge 1$ and $\omega\in\{0,1\}^m$. The depth-$m$ Sturmian partition from Section~\ref{sec:readout:empirical} has nonempty atoms, so there exists at least one word $x=x_0\cdots x_{m-1}$ in the support of $B_0\cdots B_{m-1}$ with
\[
\PP(B_0\cdots B_{m-1}=x)>0.
\]
Conditioning on that event gives
\[
\PP(A_0\cdots A_{m-1}=\omega)
\ge
\PP(B_0\cdots B_{m-1}=x)\prod_{t=0}^{m-1}Q_{x_t\omega_t},
\]
and the right-hand side is strictly positive because every channel entry is positive.
\end{proof}

\begin{proposition}[Detector realization of strict positivity]\label{prop:detector-realization}
Let $B=(B_t)_{t\in\NN}$ be the ideal Fibonacci/Sturmian benchmark process. Fix an interaction radius $L\ge 0$, coefficients $\rho_{-L},\dots,\rho_L\ge 0$ with $\sum_{j=-L}^L\rho_j>0$, a threshold $\tau\in\RR$, and i.i.d.\ detector noises $(\xi_t)_{t\in\NN}$ independent of $B$, whose common law has a density that is strictly positive on $\RR$. Define the finite-resolution detector signal and binary output by
\[
S_t:=\sum_{j=-L}^L \rho_j B_{t+j},
\qquad
A_t:=\mathbf{1}_{\{S_t+\xi_t\ge \tau\}}.
\]
Then, conditional on any neighborhood block $b_{-L}\cdots b_{m-1+L}\in\{0,1\}^{m+2L}$, the output block law factorizes as
\begin{equation*}
\begin{aligned}
\PP(&A_0\cdots A_{m-1}=a_0\cdots a_{m-1}
\mid B_{-L}\cdots B_{m-1+L}=b_{-L}\cdots b_{m-1+L})
\\
&=
\prod_{t=0}^{m-1} p_t(b)^{a_t}(1-p_t(b))^{1-a_t}.
\end{aligned}
\end{equation*}
for some probabilities $p_t(b)\in(0,1)$. In particular the detector induces a strictly positive local binary channel in the sense of Theorem~\ref{thm:noise-full-support}, so every word $\omega\in\{0,1\}^m$ has positive probability under the measured output process.
\end{proposition}

\begin{proof}
Fix a neighborhood block $b:=b_{-L}\cdots b_{m-1+L}$. For each $t=0,\dots,m-1$, the analog signal
\[
s_t(b):=\sum_{j=-L}^L \rho_j b_{t+j}
\]
is a deterministic real number. Since the common noise law of $\xi_t$ has a density that is strictly positive on $\RR$, the tail probability
\[
p_t(b):=\PP(\xi_t\ge \tau-s_t(b))
\]
lies strictly between $0$ and $1$. Independence of the detector noises therefore gives the displayed product formula.

Now fix $\omega\in\{0,1\}^m$. Choose any neighborhood block $b$ with
\[
\PP(B_{-L}\cdots B_{m-1+L}=b)>0;
\]
such a block exists because the support of the benchmark process is nonempty. Conditioning on that event and using the product formula yields
\[
\PP(A_0\cdots A_{m-1}=\omega)
\ge
\PP(B_{-L}\cdots B_{m-1+L}=b)\prod_{t=0}^{m-1} p_t(b)^{\omega_t}(1-p_t(b))^{1-\omega_t}>0.
\]
Thus the detector realization induces a strictly positive local binary channel and fills the ambient cube.
\end{proof}

\begin{remark}[Why the ambient extension is physically forced]
Theorem~\ref{thm:noise-full-support} upgrades the motivational role of the ambient binary extension to an observable-support statement, and Proposition~\ref{prop:detector-realization} realizes its hypothesis with a concrete finite-resolution detector. Once the readout layer has nonzero local uncertainty, the experimental length-$m$ support leaves the Sturmian language and fills the full cube. The stabilization problem is therefore naturally posed on $\{0,1\}^m$ in the presence of finite-resolution readout noise.
\end{remark}

\subsection{Empirical block laws for the benchmark}\label{sec:readout:empirical}

For each depth $m$, the induced law $\nu_R$ determines a raw-word marginal on $\{0,1\}^m$. In the irrational-rotation benchmark, these marginals arise from the partition of $\TT^1$ into nonempty depth-$m$ Sturmian atoms, whose number is at most $m+1$. This linear complexity is the only source-specific input needed in Section~\ref{sec:fp:tv}.

