\section{Discussion}\label{sec:discussion}

The main result is a sharp overlap threshold for a locally many-to-one Fibonacci folding. The word map $\Fold_m$ collapses residue classes on isolated windows, but Theorem~\ref{thm:block-invertibility} and Theorem~\ref{thm:two-sided-conjugacy} show that once overlapping windows are imposed, every aperture $m\ge 3$ becomes a finite-memory conjugacy onto a finite-type image.

The failure at $m=2$ is equally sharp. Rather than a diffuse non-injective regime, one finds a single branch locus: Theorem~\ref{thm:m2-branch-locus} identifies the unique exceptional orbit, Corollary~\ref{cor:m2-defect-formulas} computes the exact entropy and Kullback--Leibler defect concentrated there, and Corollary~\ref{cor:m2-mod-zero} shows that outside this branch set the two-window map is already invertible modulo null sets.

These results clarify the status of the entropy ledger. Section~\ref{sec:entropy-ledger} quantifies local aliasing of fixed windows, whereas Section~\ref{sec:measurement} proves that above threshold no new process-level irreversibility is created. The finite-memory conjugacy transports invariant measures, pressure, and equilibrium states exactly, so the stabilization image is mathematically a recoding of the full shift rather than a new dynamical system with altered thermodynamic content.

The residue formulas, discrete Fourier representation, second-moment recurrence, and discrepancy bounds then appear in the right role: they are quantitative corollaries of the folding arithmetic and the conjugacy, not substitutes for the main theorem. In the Sturmian benchmark they yield explicit deterministic control of stabilized block laws and of transported local observables.

From the sofic-shift perspective, the threshold theorem can be read as follows: the sliding stabilization map $\Phi_m$ is a right-resolving factor of degree at most $2^{m-1}$ for every $m\ge 2$. General theory guarantees that a right-resolving one-to-one factor between irreducible shifts of finite type is a conjugacy \cite{Ashley1988,LindMarcus1995}. The non-trivial content is that degree one is achieved at exactly $m=3$, and that the obstruction at $m=2$ is confined to a single orbit pair rather than a diffuse set. Neither conclusion follows from the abstract framework; both require the explicit Fibonacci residue arithmetic of Lemma~\ref{lem:m3-decoder} and Theorem~\ref{thm:m2-branch-locus}.

The Fibonacci setting is used for explicit solvability. It is the smallest aperiodic example in which the threshold decoder, the branch classification, and the thermodynamic transport can all be proved completely. Extending the same picture to more general Ostrowski or model-set numerations would require new base-case decoders and new overlap arithmetic, so that problem remains genuinely open. A natural next case is the family of metallic-mean stabilizations, where the forbidden word changes from $11$ to $1\cdots 1$ of length $k+1$; the modular arithmetic changes qualitatively and the threshold window is expected to depend on $k$.

The readout motivation from quasiperiodic order remains useful at the boundary of the paper: it explains why one is naturally led to a local folding defined on the full cube and why finite-sample discrepancy bounds matter. But the theorem-level core established here is a symbolic-dynamical and arithmetic statement about finite-memory recoding.
