\begin{thebibliography}{32}

\bibitem{Shechtman1984}
Shechtman D, Blech I, Gratias D and Cahn J W 1984
Metallic phase with long-range orientational order and no translational symmetry
\textit{Phys.\ Rev.\ Lett.} \textbf{53} 1951--3

\bibitem{Jagannathan2021}
Jagannathan A 2021
The Fibonacci quasicrystal: case study of hidden dimensions and multifractality
\textit{Rev.\ Mod.\ Phys.} \textbf{93} 045001

\bibitem{Damanik2016}
Damanik D 2016
Schr\"odinger operators with dynamically defined potentials
\textit{Ergod.\ Theory Dyn.\ Syst.} \textbf{37} 1681--764

\bibitem{Meyer1972}
Meyer Y 1972
\textit{Algebraic Numbers and Harmonic Analysis}
(Amsterdam: North-Holland)

\bibitem{Moody2000}
Moody R V 2000
Model sets: a survey
\textit{From Quasicrystals to More Complex Systems}
ed F Axel, F D\'enoyer and J-P Gazeau
(Berlin: Springer) pp~145--66

\bibitem{Moody1997}
Moody R V 1997
Meyer sets and their duals
\textit{The Mathematics of Long-Range Aperiodic Order}
ed R V Moody
(Dordrecht: Kluwer) pp~403--41

\bibitem{BaakeGrimm2013}
Baake M and Grimm U 2013
\textit{Aperiodic Order, Vol.~1: A Mathematical Invitation}
(Cambridge: Cambridge University Press)

\bibitem{Schlottmann2000}
Schlottmann M 2000
Generalized model sets and dynamical systems
\textit{Directions in Mathematical Quasicrystals}
ed M Baake and R V Moody
(Providence, RI: American Mathematical Society) pp~143--59

\bibitem{LagariasPleasants2003}
Lagarias J C and Pleasants P A B 2003
Repetitive Delone sets and quasicrystals
\textit{Ergod.\ Theory Dyn.\ Syst.} \textbf{23} 831--67

\bibitem{DrmotaTichy1997}
Drmota M and Tichy R F 1997
\textit{Sequences, Discrepancies and Applications}
(Berlin: Springer)

\bibitem{Cassels1957}
Cassels J W S 1957
\textit{An Introduction to Diophantine Approximation}
(Cambridge: Cambridge University Press)

\bibitem{CoverThomas2006}
Cover T M and Thomas J A 2006
\textit{Elements of Information Theory} 2nd edn
(Hoboken, NJ: Wiley-Interscience)

\bibitem{Kitchens1998}
Kitchens B P 1998
\textit{Symbolic Dynamics: One-Sided, Two-Sided and Countable State Markov Shifts}
(Berlin: Springer)

\bibitem{Nasu1995}
Nasu M 1995
Textile systems for endomorphisms and automorphisms of the shift
\textit{Memoirs of the American Mathematical Society} vol~114 no~546
(Providence, RI: American Mathematical Society)

\bibitem{KuipersNiederreiter1974}
Kuipers L and Niederreiter H 1974
\textit{Uniform Distribution of Sequences}
(New York: Wiley)

\bibitem{MorseHedlund1940}
Morse M and Hedlund G A 1940
Symbolic dynamics II.\ Sturmian trajectories
\textit{Am.\ J.\ Math.} \textbf{62} 1--42

\bibitem{LindMarcus1995}
Lind D and Marcus B 1995
\textit{An Introduction to Symbolic Dynamics and Coding}
(Cambridge: Cambridge University Press)

\bibitem{Frougny1992}
Frougny C 1992
Representations of numbers and finite automata
\textit{Math.\ Systems Theory} \textbf{25} 37--60

\bibitem{Berstel2001}
Berstel J 2001
An exercise on Fibonacci representations
\textit{RAIRO Theor.\ Inform.\ Appl.} \textbf{35} 491--8

\bibitem{Lothaire2002}
Lothaire M 2002
\textit{Algebraic Combinatorics on Words}
(Cambridge: Cambridge University Press)

\bibitem{AlloucheShallit2003}
Allouche J-P and Shallit J 2003
\textit{Automatic Sequences: Theory, Applications, Generalizations}
(Cambridge: Cambridge University Press)

\bibitem{Mousavi2016}
Mousavi H 2016
Automatic theorem proving in Walnut
\textit{arXiv:1603.06017 [cs.FL]}

\bibitem{DuMousaviSchaefferShallit2016}
Du C F, Mousavi H, Schaeffer L and Shallit J 2016
Decision algorithms for Fibonacci-automatic words, I: basic results
\textit{RAIRO Theor.\ Inform.\ Appl.} \textbf{50} 39--66

\bibitem{Queffelec2010}
Queff\'elec M 2010
\textit{Substitution Dynamical Systems---Spectral Analysis} 2nd edn
(Berlin: Springer)

\bibitem{Fogg2002}
Pytheas Fogg N, Berth\'e V, Ferenczi S, Mauduit C and Siegel A (ed) 2002
\textit{Substitutions in Dynamics, Arithmetics and Combinatorics}
(Berlin: Springer)

\bibitem{Zeckendorf1972}
Zeckendorf E 1972
Repr\'esentation des nombres naturels par une somme de nombres de Fibonacci ou de nombres de Lucas
\textit{Bull.\ Soc.\ Roy.\ Sci.\ Li\`ege} \textbf{41} 179--82

\bibitem{Ostrowski1922}
Ostrowski A 1922
Bemerkungen zur Theorie der diophantischen Approximationen
\textit{Abh.\ Math.\ Semin.\ Univ.\ Hambg.} \textbf{1} 77--98

\bibitem{Berthe2001Ostrowski}
Berth\'e V 2001
Autour du syst\`eme de num\'eration d'Ostrowski
\textit{Bull.\ Belg.\ Math.\ Soc.\ Simon Stevin} \textbf{8} 209--39

\bibitem{Walters1982}
Walters P 1982
\textit{An Introduction to Ergodic Theory}
(New York: Springer)

\bibitem{Sharma2007}
Sharma H R, Shimoda M and Tsai A P 2007
Quasicrystal surfaces: structure and growth of atomic overlayers
\textit{Adv.\ Phys.} \textbf{56} 403--64

\bibitem{Erdos1939}
Erd\H{o}s P 1939
On a family of symmetric Bernoulli convolutions
\textit{Amer.\ J.\ Math.} \textbf{61} 974--6

\bibitem{Garsia1963}
Garsia A M 1963
Entropy and singularity of infinite convolutions
\textit{Pacific J.\ Math.} \textbf{13} 1159--69

\bibitem{Solomyak1995}
Solomyak B 1995
On the random series $\sum\pm\lambda^n$ (an Erd\H{o}s problem)
\textit{Ann.\ of Math.} \textbf{142} 611--25

\bibitem{Parry1960}
Parry W 1960
On the $\beta$-expansions of real numbers
\textit{Acta Math.\ Acad.\ Sci.\ Hungar.} \textbf{11} 401--16

\bibitem{BertheRigo2010}
Berth\'e V and Rigo M (ed) 2010
\textit{Combinatorics, Automata and Number Theory}
(Encyclopedia of Mathematics and its Applications vol~135)
(Cambridge: Cambridge University Press)

\bibitem{Ashley1988}
Ashley J 1988
Resolving factor maps for shifts of finite type with equal entropy
\textit{Ergod.\ Theory Dyn.\ Syst.} \textbf{8} 367--78

\bibitem{Boyle1983}
Boyle M 1983
Lower entropy factors of sofic systems
\textit{Ergod.\ Theory Dyn.\ Syst.} \textbf{3} 541--57

\bibitem{BertheSteinerThuswaldner2019}
Berth\'e V, Steiner W and Thuswaldner J M 2019
Geometry, dynamics, and arithmetic of $S$-adic shifts
\textit{Ann.\ Inst.\ Fourier} \textbf{69} 1347--409

\end{thebibliography}
